```{r setup, echo=FALSE, include=FALSE}
# ############################################
# Generate tables and figures for a periodic TB report
# Tom Hiatt
# Updated: 10 Apr 2014
# ############################################

# Define structure of data (customize as needed)
# ----------------------------

# Period to examine and final
prd.group <- "year"
prd <- 2013

# Upper and lower aggregate levels
upper.var <- "all"
upper.label <- "the Philippines"
lower.var <- "region"
lower.label <- "Region"

# Import data (customize as needed)
# ---------------------------
tb1 <- read.csv("D:/Users/hiattt/Dropbox/STB-PHL/Annual report PHL/PHLraw2.csv") 

if(upper.var != "all") tb1 <- subset(tb1, upper.var==upper.label)

# fix factors to be in order of data
# [also replace with something elegant]
for(var in 1:ncol(tb1)){
  if(is.factor(tb1[,var])) tb1[,var] <- factor(tb1[,var], levels=unique(tb1[,var]))
}

# Aggregate data set
# ---------------------------
# [replace with something elegant]
int.vars <- NULL
for(var in 1:ncol(tb1)){
  int.vars[var] <- is.integer(tb1[,var]) | is.numeric(tb1[,var])
}

ltb <- aggregate(tb1[int.vars], by=list(area=tb1[[lower.var]], prd.group=tb1[[prd.group]]), FUN=sum, na.rm=TRUE)

utb <- aggregate(tb1[int.vars], by=list(area=rep(upper.label, nrow(tb1)), prd.group=tb1[[prd.group]]), FUN=sum, na.rm=TRUE)

tb <- rbind(ltb, utb)

# Load necessary packages
# ---------------------------
library(ggplot2)
library(reshape)
library(ggthemes)
library(scales)
library(stringr)

# Graph theme components
# ---------------------------
theme_report <- function(base_size=12, base_family="") {
    colors <- ggthemes_data$few
    gray <- colors$medium['gray']
    black <- colors$dark['black'] # I don't know why these last 3 parts are needed, but they are. List is here: http://docs.ggplot2.org/0.9.2.1/theme.html
    theme_bw(base_size=base_size, base_family=base_family) +
        theme(
            line = element_line(colour = gray),
            rect = element_rect(fill = "white", colour = NA),
            text = element_text(colour = black),
            axis.ticks.x = element_line(colour = gray),
            axis.ticks.y = element_blank(),
            legend.key = element_rect(colour = NA),
            ## Examples do not use grid lines
            panel.border = element_rect(colour = gray),
            panel.grid.major.x = element_blank(),
            panel.grid.minor = element_blank(),
            strip.background = element_rect(fill="white", colour=NA),
            strip.text = element_text(hjust=0)
            )
}

# Useful functions
# ---------------------------
library(grid)
# pos - where to add new labels
# newpage, vp - see ?print.ggplot
facetAdjust <- function(x, pos = c("up", "down"), 
                        newpage = is.null(vp), vp = NULL)
{
  # part of print.ggplot
  ggplot2:::set_last_plot(x)
  if(newpage)
    grid.newpage()
  pos <- match.arg(pos)
  p <- ggplot_build(x)
  gtable <- ggplot_gtable(p)
  # finding dimensions
  dims <- apply(p$panel$layout[2:3], 2, max)
  nrow <- dims[1]
  ncol <- dims[2]
  # number of panels in the plot
  panels <- sum(grepl("panel", names(gtable$grobs)))
  space <- ncol * nrow
  # missing panels
  n <- space - panels
  # checking whether modifications are needed
  if(panels != space){
    # indices of panels to fix
    idx <- (space - ncol - n + 1):(space - ncol)
    # copying x-axis of the last existing panel to the chosen panels 
    # in the row above
    gtable$grobs[paste0("axis_b",idx)] <- list(gtable$grobs[[paste0("axis_b",panels)]])
    if(pos == "down"){
      # if pos == down then shifting labels down to the same level as 
      # the x-axis of last panel
      rows <- grep(paste0("axis_b\\-[", idx[1], "-", idx[n], "]"), 
                   gtable$layout$name)
      lastAxis <- grep(paste0("axis_b\\-", panels), gtable$layout$name)
      gtable$layout[rows, c("t","b")] <- gtable$layout[lastAxis, c("t")]
    }
  }
  # again part of print.ggplot, plotting adjusted version
  if(is.null(vp)){
    grid.draw(gtable)
  }
  else{
    if (is.character(vp)) 
      seekViewport(vp)
    else pushViewport(vp)
    grid.draw(gtable)
    upViewport()
  }
  invisible(p)
}

```

Tuberculosis in `r upper.label`: analysis of case notification data in `r prd`
========================================================

1.  Introduction
--------------------------------------------------------

2.	Accomplishment (Drug susceptible)
--------------------------------------------------------

### a.	Case notification for all TB cases -2000 to 2013

#### i.	National - by type of TB cases (new smear positive, new smear negative, EP, relapse and re-treatment cases) annually

```{r fig-notif-trend, echo=FALSE}
notb <- subset(utb, prd.group >= 2000, c(prd.group, new_sp, new_sn, ret_rel, new_ep, new_sp, e_pop_num))

notb$tot_newrel <- notb$new_sp + notb$new_sn + notb$ret_rel + notb$new_ep 

notb$`All forms of TB` <- notb$tot_newrel / notb$e_pop_num * 1e5
notb$`New smear-positive` <- notb$new_sp / notb$e_pop_num * 1e5
notc <- melt(notb[c("prd.group", "New smear-positive", "All forms of TB")], id=1)

# write.csv(notc, file=paste0(pasteLabel("./figure data/figure", figCount, "f-notif-trend", insLink=FALSE), ".csv"), row.names=FALSE)

ggplot(notc, aes(prd.group, value, color=variable, ymin=0)) + geom_line(size=1) + theme_report() + scale_x_continuous("") + scale_y_continuous("TB cases per 100 000", breaks = pretty_breaks()) + theme(legend.position="none") + geom_text(data=subset(notc, prd.group==max(notc$prd.group)), aes(label=variable), hjust=1.1, vjust=3) + scale_color_brewer(type="qual", palette=6) + ggtitle(paste0("Tuberculosis case notification rate (all forms and new smear-positive) \nper 100 000 population in ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group)))

```

```{r fig-notif-trend-types, echo=FALSE}
notb <- subset(utb, prd.group >= 2000, c(area, prd.group, new_sp, new_sn, ret_rel, new_ep, new_sp, e_pop_num))

notb$`New smear-positive` <- notb$new_sp
notb$`New smear-negative` <- notb$new_sn
notb$`New extrapulmonary` <- notb$new_ep
notb$`Relapse` <- notb$ret_rel

notc <- melt(notb[c("area", "prd.group", "New smear-positive", "New smear-negative", "New extrapulmonary", "Relapse")], id=1:2)

ggplot(notc, aes(prd.group, value, color=variable, ymin=0)) + geom_line(size=1) + theme_report() + scale_x_continuous("") + scale_y_continuous("TB cases per 100 000", breaks = pretty_breaks()) + theme(legend.position="none") + geom_text(data=subset(notc, prd.group==max(notc$prd.group)), aes(label=variable), hjust=1.1, vjust=3) + scale_color_brewer(type="qual", palette=6) + ggtitle(paste0("Tuberculosis case notification by case type \nin ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group)))

```

#### ii.	Regional - y type of TB cases (new smear positive, new smear negative, EP, relapse and re-treatment cases) annually

```{r fig-notif-trend-type-lower, echo=FALSE}
notb <- subset(ltb, prd.group >= 2000, c(area, prd.group, new_sp, new_sn, ret_rel, new_ep, new_sp, e_pop_num))

notb$`New smear-positive` <- notb$new_sp
notb$`New smear-negative` <- notb$new_sn
notb$`New extrapulmonary` <- notb$new_ep
notb$`Relapse` <- notb$ret_rel

notc <- melt(notb[c("area", "prd.group", "New smear-positive", "New smear-negative", "New extrapulmonary", "Relapse")], id=1:2)

ggplot(notc, aes(prd.group, value, color=variable, ymin=0)) + geom_point(size=1) + geom_line(size=1)  + facet_wrap(~area, scales="fixed") + theme_report() + scale_x_continuous("") + scale_y_continuous("TB cases", breaks = pretty_breaks()) + scale_color_brewer("Case type", type="qual", palette=6) + ggtitle(paste0("Tuberculosis case notification by case type \nand ", tolower(lower.label)," in ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group))) + theme(legend.position="bottom")


```

### b.	Case notification for new smear positive by age and sex annually

#### i.	National

#### ii.  Regional

```{r fig-agesex-trend,  echo=FALSE}

aga <- subset(tb, year>1999, select=c("area", "prd.group", "new_sp_m04", "new_sp_m514", "new_sp_m014", "new_sp_m1524", "new_sp_m2534", "new_sp_m3544", "new_sp_m4554", "new_sp_m5564", "new_sp_m65", "new_sp_mu", "new_sp_f04", "new_sp_f514", "new_sp_f014", "new_sp_f1524",  "new_sp_f2534", "new_sp_f3544", "new_sp_f4554", "new_sp_f5564", "new_sp_f65", "new_sp_fu", "e_pop_m04", "e_pop_m514", "e_pop_m014", "e_pop_m1524", "e_pop_m2534", "e_pop_m3544", "e_pop_m4554", "e_pop_m5564", "e_pop_m65", "e_pop_f04", "e_pop_f514", "e_pop_f014", "e_pop_f1524", "e_pop_f2534", "e_pop_f3544", "e_pop_f4554", "e_pop_f5564", "e_pop_f65"))

# # Aggregate for region
# agb1 <- aggregate(aga[6:ncol(aga)], by=list(year=aga$year, area=aga$g_whoregion), FUN=sum, na.rm=TRUE)
# 
# agb1[c('iso3', 'g_hbc22')] <- NA
# agb1$g_whoregion <- agb1$area
# aga1 <- rename(aga,c('country'='area'))
# agb <- rbind(aga1, agb1)

# Calculate rates

agc <- melt(aga, id=1:2)

agc$type <- str_extract(agc$variable, 'new_sp|pop')
agc$age <- str_replace(agc$variable, "new_sp_|e_pop_", "")
agc$sex <- str_replace(agc$age, '[u]|[0-9]+', "")
agc$age <- str_replace(agc$age, '[f]|[m]', "")

agd <- cast(subset(agc, select= -variable), ...~type)

agd$new_sp_100k <- agd$new_sp / agd$pop * 1e5

# Subset for age groups
age <- subset(agd, age!="u" & !age %in% c("04", "514"))
age$sex <- factor(age$sex, levels=c("m", "f"), labels=c("Male", "Female"))

age$age <- factor(age$age, c("014", "1524", "2534", "3544", "4554", "5564", "65"), c("0–14", "15–24", "25–34", "35–44", "45–54", "55–64", "\u2265 65"))

ggplot(age, aes(prd.group, new_sp_100k, color=age)) + geom_line() + facet_grid(area~sex, scales='free_y') + theme_report() + scale_x_continuous("", breaks = pretty_breaks()) + scale_color_brewer("Age group", type="seq", palette=4) + scale_y_continuous("New smear-positive cases per 100 000") + ggtitle(paste0("New smear-positive TB case rates by age and sex \nand by ", tolower(lower.label)," in ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group)))

```


### c.	Casefinding

#### i.	National - Number of TB symptomatics examined by smear microscopy by year 

#### ii.	Regional - Number of TB symptomatics examined by smear microscopy by year 
(Should this be per population to make it comparable?)

```{r fig-sympt-smear-lower, echo=FALSE}

symp <- ggplot(tb, aes(prd.group, sympt_sm)) + geom_point(size=1) + geom_line(size=1)  + facet_wrap(~area, scales="free_y") + theme_report() + scale_x_continuous("") + scale_y_continuous("TB symptomatics examined by smear microscopy", breaks = pretty_breaks()) + ggtitle(paste0("TB symptomatics examined by smear microscopy by \n", tolower(lower.label)," in ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group))) 

facetAdjust(symp)

```


#### iii.	Summary of laboratory activities per year per region

1.	Number of TB symptomatics examined
2.	Number with three sputum specimens
3.	Number of smear positive
4.	Positivity rate

### d.	Caseholding

#### i.	National

1.	New smear positive
2.	New smear negative
3.	Relapse + Failure + RAD

#### ii.	National

1.	New smear positive
2.	New smear negative
3.	Relapse + Failure + RAD

3.	Accomplishment (Drug resistant) 2000 to 2013
--------------------------------------------------------

### a.	Casefinding

#### i.	National

### b.	Caseholding

#### i.	National 

4.	Accomplishment – Children
--------------------------------------------------------

5.	Accomplishment – PPMD
--------------------------------------------------------

6.	Accomplishment – Jails and Prisons
--------------------------------------------------------

7.	Accomplishment – TB/HIV
--------------------------------------------------------

8.	Policies
--------------------------------------------------------

9.	Budget
--------------------------------------------------------

10.	Partners
--------------------------------------------------------

