```{r setup, echo=FALSE, include=FALSE}
# ############################################
# Generate tables and figures for a periodic TB report
# Tom Hiatt
# Updated: 10 Apr 2014
# ############################################

# Define structure of data
# ----------------------------

# Period to examine and final
prd.group <- "year"
prd <- 2013

# Upper and lower aggregate levels
upper.var <- "all"
upper.label <- "the Philippines"
lower.var <- "region"
lower.label <- "Region"

# Import data
# ---------------------------
tb1 <- read.csv("D:/Users/hiattt/Dropbox/STB-PHL/Annual report PHL/PHLraw.csv") 

if(upper.var != "all") tb1 <- subset(tb1, upper.var==upper.label)

# Aggregate data set
# ---------------------------
# [replace with something elegant]
int.vars <- NULL
for(var in 1:ncol(tb1)){
  int.vars[var] <- is.integer(tb1[,var]) | is.numeric(tb1[,var])
}

ltb <- aggregate(tb1[int.vars], by=list(area=tb1[[lower.var]], prd.group=tb1[[prd.group]]), FUN=sum, na.rm=TRUE)

utb <- aggregate(tb1[int.vars], by=list(area=rep(upper.label, nrow(tb1)), prd.group=tb1[[prd.group]]), FUN=sum, na.rm=TRUE)

tb <- rbind(ltb, utb)

# Load necessary packages
# ---------------------------
library(ggplot2)
library(reshape)
library(ggthemes)
library(scales)

# Graph theme components
# ---------------------------
theme_report <- function(base_size=12, base_family="") {
    colors <- ggthemes_data$few
    gray <- colors$medium['gray']
    black <- colors$dark['black'] # I don't know why these last 3 parts are needed, but they are. List is here: http://docs.ggplot2.org/0.9.2.1/theme.html
    theme_bw(base_size=base_size, base_family=base_family) +
        theme(
            line = element_line(colour = gray),
            rect = element_rect(fill = "white", colour = NA),
            text = element_text(colour = black),
            axis.ticks.x = element_line(colour = gray),
            axis.ticks.y = element_blank(),
            legend.key = element_rect(colour = NA),
            ## Examples do not use grid lines
            panel.border = element_rect(colour = gray),
            panel.grid.major.x = element_blank(),
            panel.grid.minor = element_blank(),
            strip.background = element_rect(fill="white", colour=NA),
            strip.text = element_text(hjust=0)
            )
}



```

Tuberculosis in `r upper.label`: analysis of case notification data in `r prd`
========================================================

1.  Introduction
--------------------------------------------------------

2.	Accomplishment (Drug susceptible)
--------------------------------------------------------

### a.	Case notification for all TB cases -2000 to 2013

#### i.	National - by type of TB cases (new smear positive, new smear negative, EP, relapse and re-treatment cases) annually

```{r f-notif-trend, echo=FALSE}
notb <- subset(utb, prd.group >= 2000, c(prd.group, new_sp, new_sn, ret_rel, new_ep, new_sp, e_pop_num))

notb$tot_newrel <- notb$new_sp + notb$new_sn + notb$ret_rel + notb$new_ep 

notb$`All forms of TB` <- notb$tot_newrel / notb$e_pop_num * 1e5
notb$`New smear-positive` <- notb$new_sp / notb$e_pop_num * 1e5
notc <- melt(notb[c("prd.group", "New smear-positive", "All forms of TB")], id=1)

# write.csv(notc, file=paste0(pasteLabel("./figure data/figure", figCount, "f-notif-trend", insLink=FALSE), ".csv"), row.names=FALSE)

ggplot(notc, aes(prd.group, value, color=variable, ymin=0)) + geom_line(size=1) + theme_report() + scale_x_continuous("") + scale_y_continuous("TB cases per 100 000", breaks = pretty_breaks()) + theme(legend.position="none") + geom_text(data=subset(notc, prd.group==max(notc$prd.group)), aes(label=variable), hjust=1.1, vjust=3) + scale_color_brewer(type="qual", palette=6) + ggtitle(paste0("Tuberculosis case notification rate (all forms and new smear-positive) \nper 100 000 population in ", upper.label, ", ", min(notc$prd.group), "–", max(notc$prd.group)))

```

#### ii.	Regional - y type of TB cases (new smear positive, new smear negative, EP, relapse and re-treatment cases) annually

```{r f-notif-trend-lower, echo=FALSE}
notd <- subset(ltb, prd.group >= 2000, c(area, prd.group, new_sp, new_sn, ret_rel, new_ep, new_sp, e_pop_num))

notd$tot_newrel <- notd$new_sp + notd$new_sn + notd$ret_rel + notd$new_ep 

notd$`All forms of TB` <- notd$tot_newrel / notd$e_pop_num * 1e5
notd$`New smear-positive` <- notd$new_sp / notd$e_pop_num * 1e5
note <- melt(notd[c("area", "prd.group", "New smear-positive", "All forms of TB")], id=1)

# write.csv(note, file=paste0(pasteLabel("./figure data/figure", figCount, "f-notif-trend", insLink=FALSE), ".csv"), row.names=FALSE)

ggplot(note, aes(prd.group, value, color=variable, ymin=0)) + geom_line(size=1) + theme_report() + scale_x_continuous("") + scale_y_continuous("TB cases per 100 000", breaks = pretty_breaks()) + theme(legend.position="none") + geom_text(data=subset(note, prd.group==max(note$prd.group)), aes(label=variable), hjust=1.1, vjust=3) + scale_color_brewer(type="qual", palette=6) + ggtitle(paste0("Tuberculosis case notification rate (all forms and new smear-positive) \nper 100 000 population in ", upper.label, ", ", min(note$prd.group), "–", max(note$prd.group)))

```

### b.	Case notification for new smear positive by age and sex annually

#### i.	National

#### ii.	Regional

### c.	Casefinding

#### i.	National - Number of TB symptomatics examined by smear microscopy by year 

#### ii.	Regional - Number of TB symptomatics examined by smear microscopy by year 

#### iii.	Summary of laboratory activities per year per region

1.	Number of TB symptomatics examined
2.	Number with three sputum specimens
3.	Number of smear positive
4.	Positivity rate

### d.	Caseholding

#### i.	National

1.	New smear positive
2.	New smear negative
3.	Relapse + Failure + RAD

#### ii.	National

1.	New smear positive
2.	New smear negative
3.	Relapse + Failure + RAD

3.	Accomplishment (Drug resistant) 2000 to 2013
--------------------------------------------------------

### a.	Casefinding

#### i.	National

### b.	Caseholding

#### i.	National 

4.	Accomplishment – Children
--------------------------------------------------------

5.	Accomplishment – PPMD
--------------------------------------------------------

6.	Accomplishment – Jails and Prisons
--------------------------------------------------------

7.	Accomplishment – TB/HIV
--------------------------------------------------------

8.	Policies
--------------------------------------------------------

9.	Budget
--------------------------------------------------------

10.	Partners
--------------------------------------------------------

