Regional TB report `r yr <- as.numeric(format(Sys.time(),"%Y")) - 1; yr+1`
========================================================

Abstract
--------------------------------------------------------

### Background

### Methods

### Results

### Conclusion

Introduction
--------------------------------------------------------

This report is the first in a series of reports.

Methods
--------------------------------------------------------

Data were collected through an on-line data collection system.

Results
--------------------------------------------------------

```{r echo=FALSE}
# runprofile()
# Functions to assist this document
incCount <- function(inObj, useName) {
  nObj <- length(inObj)
  useNum <- max(inObj) + 1
  inObj <- c(inObj, useNum)
  names(inObj)[nObj + 1] <- useName
  inObj
}
figCount <- c(`_` = 0)
tableCount <- c(`_` = 0)

# tableCount <- incCount(tableCount, "t.blogPostDocs")
# tableCount

pasteLabel <- function(preText, inObj, objName, insLink = TRUE) {
  objNum <- inObj[objName]
  
  useText <- paste(preText, objNum, sep = " ")
  if (insLink) {
    useText <- paste("[", useText, "](#", objName, ")", sep = "")
  }
  useText
}

tableCat <- function(inFrame) {
    outText <- paste(names(inFrame), collapse = " | ")
    outText <- c(outText, paste(rep("---", ncol(inFrame)), collapse = " | "))
    invisible(apply(inFrame, 1, function(inRow) {
        outText <<- c(outText, paste(inRow, collapse = " | "))
    }))
    return(outText)
}
```

``` {r, echo=FALSE}
# Order list
figCount <- incCount(figCount, "f.blogPostDocs")
tableCount <- incCount(tableCount, "t.blogPostDocs")
```

As seen in `r I(pasteLabel("Figure", figCount, "f.blogPostDocs"))` the trend is evident.

<a id="f.blogPostDocs"></a> 
```{r fig.width=7, fig.height=6, echo=FALSE}

plot(cars)
```
#### `r I(pasteLabel("Figure", figCount, "f.blogPostDocs", insLink=FALSE))` Case rates per 100 000 population, `r yr`.

And also as seen in `r I(pasteLabel("Table", tableCount, "t.blogPostDocs"))` the trend is evident.

<a id="t.blogPostDocs"></a> 
```{r echo=FALSE, results='asis'}
postDocBlogs <- data.frame(PD = c("p1", "p2", "p3"), NBlog = c(4, 10, 2), Year = c(1, 
    4, 2))

postDocInsert <- tableCat(postDocBlogs)
```

``` {r echo=FALSE, results='asis'}
cat(postDocInsert, sep = "\n")
```
#### `r I(pasteLabel("Table", tableCount, "t.blogPostDocs", insLink=FALSE))` Case rates per 100 000 population, `r yr`.

Overall header

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell


Discussion
--------------------------------------------------------

Annex
--------------------------------------------------------

The

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

```{r}

```
