---
title: "Drug resistant tuberculosis in the WHO Western Pacific Region"
author: "Tauhid Islam,<sup>&dagger;</sup> Tom Hiatt,<sup>&dagger;</sup> Cornelia Hennig,<sup>&dagger;</sup>
  Nobuyuki Nishikiori<sup>&dagger;</sup>"
csl: springer-vancouver.csl
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: no
    theme: united
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    fig_height: 7
bibliography: MDRupdate.bib
---


```{r setup, echo=FALSE, include=FALSE}
# THE PARTS BETWEEN ``` ARE R CODE AND CAN BE IGNORED. SEARCH FOR THE SECTION YOU WANT TO EDIT (EG “ABSTRACT”)
# Script found at: https://docs.google.com/document/d/1QGiWYD68Y6w1lo8d2or4889kSxHrzcNkMK3lhPQOvCQ/edit

source("./MDRupdate.R") # change before publication
# source("./script2.R")

```




<sup>&dagger;</sup> Stop TB and Leprosy Elimination, Division of Communicable Disease, World Health Organization Regional Office for the Western Pacific, Manila, Philippines

Running title: DR-TB in the WHO Western Pacific Region

Key words: tuberculosis; surveillance; drug-resistance; MDR-TB; Western Pacific Region

Corresponding author: 
Tauhid Islam
Stop TB and Leprosy Elimination
Division of Communicable Disease
World Health Organization Regional Office for the Western Pacific
Manila, Philippines 
Tel: +632 5289720
Email: islamt@wpro.who.int


Abstract 
--------------------------------------------------------

Objective: To review the latest information about levels of anti TB drug resistant and various indicators of programmatic management of drug resistant TB in the Western Pacific Region of the World Health Organization (WHO).

Methods: we analysed routine data collected by WHO annually……

Findings: 4% (95% confidence limit 3-6) of new and 22% (95% confidence limit 18-27)  of previously treated TB cases are estimated to have MDR-TB in the Region. DST coverage among new is 3% and among retreatment cases is 21%.XXX cases are notified in 2013 which is 15% of the estimated MDR-TB cases. Among the notified cases, XXX or 65% are put on treatment.  AMong all MDR-TB cases 38% had DST to second line drug sand 7% are resistant to either or both FQ and injectable agent. Treatment outcome of 2011 cohort showed 50% success rate, X failed and X% loss to follow up.Over last 5 years case notification and enrolment has increased but also gap between notification and enrolment incresaed.

Discussion:  The increasing trend in detection and enrolment of MDR-TB cases demonstrates readiness for PMDT scale-up at the country level. However, challenges remain considerable. More complete data will be helpful to understand the situation and address the gaps. 


Introduction
--------------------------------------------------------

Even if less than 5% of the TB burden is MDR-TB, this poses a formidable challenge to the prospect of controlling TB for the following reasons: (1) as incidence of TB declines, drug resistant strains may dominate the epidemic (2) treatment duration is too long, around 2 years, and complicated due to severe side effects of second-line drugs (3) management costs are too high and responsible for catastrophic patient expenditure (4) treatment outcome is very poor and (5) it poses a huge burden on existing health systems as human resource and technical capacity are not ready to tackle this issue. 

Globally, a total of 300 000 (range 230 000–380 000) people are estimated to have developed multidrug-resistant TB (MDR-TB, resistant to at least rifampicin and isoniazid) in `r yr` [@WorldHealthOrganization2014]. Among them, approximately one quarter MDR-TB occurred in countries of the Western Pacific Region of the World Health Organization (WHO). Most (more than `r tc <- subset(mdat, g_whoregion=="WPR" & year==yr, c(iso3, e_mdr_num)); round(sum(tc[tc$iso3 %in% c("CHN", "PHL", "VNM"), "e_mdr_num"])/sum(tc$e_mdr_num, na.rm=TRUE) * 100, 0)`%) of the MDR-TB patients are in three countries, namely China, the Philippines and Viet Nam.  

The Region has made substantial progress and reached Millennium Development Goal (MDG) targets for TB control in advance of 2015.  TB prevalence and mortality are below half of 1990 levels, and case detection and treatment success remain high (Reference
…).  However, the challenges of TB remain monumental.  Annually, an estimated `r subset(madat, group_name=="WPR" & year==yr, e_inc_num) / 1e6` million people develop TB, and `r subset(madat, group_name=="WPR" & year==yr, e_mort_exc_tbhiv_num)/1000` thousand die from this curable illness.  TB concentrates in vulnerable populations, such as migrants, children, elderly and poor people.  Meanwhile, every year the Region’s MDR-TB burden increases by an estimated 71 thousand cases---only 9.8% of which are treated under quality-assured programmes.  

To address the problem of drug resistance in the WHO Western Pacific Region, the Regional strategy to stop tuberculosis in the Western Pacific 2011--2015 identified scaling up the programmatic management of drug resistant TB as one of the five objectives.[@WorldHealthOrganization2011] Surveillance of resistance to anti-TB drugs, well balanced capacity of diagnosing, treating, availability of quality assured drugs  and monitoring the progress of programmatic implementation are fundamental components of the strategy. In 2011, the Western Pacific Region Green Light Committee (WPR rGLC) was established to assist member countries with responsible PMDT scale-up. 

This article is the second in a series of regional TB situation analysis after ‘epidemiology and control of tuberculosis in the Western Pacific Region: analysis of 2012 case notification data’.[@Hiatt2014] In this article, we review the latest available information about drug resistance in the WHO Western Pacific Region and its programmatic implementation. We hope that a thorough analysis of case detection, notification, enrolment and outcome data provides valuable information on programmatic progress and future direction. We also provide a brief overview of WPR rGLC activities, reflect on achievements and challenges, and explore its potential future role.




Methods
--------------------------------------------------------

The WHO Western Pacific Region includes `r te <- subset(mdat, g_whoregion=="WPR" & year==yr, e_pop_num); nrow(te)` countries and areas.  In `r yr`, `r tf <- subset(mdat, g_whoregion=="WPR" & year==yr & !is.na(c_notified), e_pop_num); nrow(tf)` countries and areas of the Western Pacific Region reported data representing more than `r signif(sum(tf)/sum(te)*100, 4)`% of the total population. Since 2009, a web-based online system has been used for data submission and validation. Collected data covers the following areas: TB case notifications and treatment outcomes, diagnostic and treatment services, drug management, surveillance and surveys of drug-resistance, information on TB/HIV co-infection, infection control, engagement of all care providers and budgets and expenditures for TB control. The full description of methods is available in the Global Tuberculosis Report `r yr` and the data sets are available from the WHO global TB database (www.who.int/tb/data). This report described the drug resistance situation and progress in programmatic response in all countries of the Region where at least 1 MDR-TB case was detected and a special focus was maintained on seven countries that are considered as priority countries with a high burden of TB in the Western Pacific Region: Cambodia, China, The Lao People’s Democratic Republic, Mongolia, Papua New Guinea, the Philippines and Viet Nam.

### Analysis and reproducibility
```{r Versions}
info <- sessionInfo()
r_ver <- paste(info$R.version$major, info$R.version$minor, sep=".")
```
 
Analysis was conducted by the statistical package R ((ver. `r r_ver`), R Core Team, 2013, Vienna, Austria, www.R-project.org). Due to calls for transparent and reproducible research, we have published programme code to generate the entire contents of this article including all figures and tables by using R with the knitr package (ver. `r info$otherPkgs$knitr$Version`, Yihui Xie, 2013). Readers can download the code (see supplement material) and reproduce all figures and tables under an appropriate personal computing environment. For non-commercial purposes, readers may modify the code to produce figures and tables that are not presented in this article. For instance, readers may wish to produce tables and figures for countries or regions other than the WHO Western Pacific Region.

```{r results='asis', echo=FALSE, message=FALSE}
# setDownloadURI(list("tb", "script1", "script2"), filename = "MDRWPRO2014", textHTML = "Click here to download the data and scripts.")
```


Findings
--------------------------------------------------------

### Coverage of drug resistance surveillance 

Estimate of MDR-TB is based on the availability of drug resistance data either from surveillance or from special survey. Drug resistance data were available from `r tg <- subset(mdat, g_whoregion=="WPR" & year==yr & !is.na(mdr), e_pop_num); nrow(tg)` of the `r nrow(te)` countries and areas of the WHO Western Pacific Region and `r length(whbc)` countries reported at least one MDR-TB case. Among the countries where drug resistance data are available, 19 countries rely on surveillance data and 6 countries rely on special survey of representative samples of patients. Nationwide and population representative data were available from 5 countries (China, Cambodia, Mongolia, Philippines and Viet Nam). Malaysia has data from certain geographical area that are not considered representative of the country a whole and the data is from before 2000. Among high TB burden countries, Lao PDR and PNG do not have drug resistance data available yet and the estimate of MDR-TB is based on extrapolation of the Regional figure. Drug resistance surveys is ongoing in Papua New Guinea which is expected to produce sub national data in 2015. 

<a id="m.coverage"></a> 

##### `r I(pasteLabel("Figure", figCount, "m.coverage", insLink=FALSE))`.  Progress in regional coverage of surveillance data on drug resistance, `r paste0(range(coa$year), collapse="\u2013")`^a^
``` {r m.coverage, fig.width=9, fig.height=9}
m.coverage
```

^a^ Coverage in Malaysia and Papua New Guinea is at the subnational level only.


<a id="t.drestnotif"></a> 

##### `r I(pasteLabel("Table", tableCount, "t.drestnotif", insLink=FALSE))`. Estimated drug resistant cases, `r yr`
```{r t-drestnotif, results='asis'}
t.drestnotif
```


### MDR-TB rate
The Region includes Member States with a wide range of MDR-TB rate and case load. Table 1 shows the rate of MDR-TB of countries and areas reported at least 1 MDR-Tb case in 2013. The proportion of previously treated TB cases with MDR-TB ranged from 0% to 34%. Mongolia has the highest estimated proportion 34% (95% CI: 29-38) of MDR-TB among retreatment cases. Countries with more than 20% MDR-TB among retreatment cases are China (26%), Mongolia (34%), New Zealand(20%), Philippines (21%) and Viet Nam (23%).The proportion of new TB cases with MDR-TB ranged from 0% to 6%.  China has the highest proportion 6% (95% CI: 5-7) of MDR-TB among new cases.Combining data from all countries of the Region, 4% (95% confidence interval: 3-6) of new and 22% (95% CI: 18-26) of previously treated TB cases were estimated to have MDR-TB.  Though highest % of MDR-TB is observed among retreatment cases, absolute number of estimated MDR cases is lower among retreatment cases than the new cases. Among the total estimated MDR-TB, 24% are from retreatment cases and 76% are from new cases.




### Management of drug resistant TB: 
  	
#### DST coverage


Drug sensitivity testing data was reported from 19 countries (53%  of 36 Member Countries and Areas). Only 3% of new bacteriologically confirmed TB cases and 21% of those previously treated for TB were tested for MDR-TB or Rifampicin resistance status. Among  previously treated cases, the coverage of testing remain below 50% in China (20%), japan (43%), Lao PDR (26%), Malaysia (9%), Marshall Island (43%), Philippines (14%) and Republic of Korea (9%). Testing for MDR-TB among new cases, the coverage was less than 10% in China (3%), Lao PDR (<1%), Philippines (<1%), Republic of Korea (4%) and Viet Nam (3%). The testing coverage among retreatment is increasing steadily over the years while MDR-TB positivity rate also remained high.

<a id="f.dst.trend"></a> 

##### `r I(pasteLabel("Figure", figCount, "f.dst.trend", insLink=FALSE))`. Trend of drug-resistance coverage and positivity of MDR-TB and RR-TB amongst retreatment patients, `r paste(2007, yr, sep="\u2013")`
```{r f.dst.trend, fig.width=9, fig.height=9}
f.dst.trend
```

^a^ Drug-resistance testing includes testing by Xpert for Rifampicin resistance only.

Among cases with phenotypic DST result 8% has resistance to isoniazid (H) only, 2% has resistant to Rifampicin (R) only and 11% has resistant both to H and R (MDR-TB).

Countries started using the rapid diagnostic technology,Xpert MTB/Rif which was endorsed by WHO in 2010 (Reference) and R resistant cases are being put on MDR-Tb treatment. In 2013, additional 17% cases has been identified by Xpert MTB/Rif test. In Viet Nam 81%, in Papua New guinea and 72% MDR-TB cases are identified by Xpert MTB/Rif. 

<a id="t.drnotif"></a> 

##### `r I(pasteLabel("Table", tableCount, "t.drnotif", insLink=FALSE))`. Drug resistance in select countries, `r yr`
```{r t.drnotif, results='asis'}
t.drnotif
```


#### Notification and enrolment


<a id="t.est.enroll"></a> 

##### `r I(pasteLabel("Table", tableCount, "t.est.enroll", insLink=FALSE))`. Estimated drug resistant cases compared with notified cases and cases put on treatment, `r yr`
```{r t.est.enroll, results='asis'}
t.est.enroll
```


The 10578 reported cases of Rif resistant (considered as MDR-TB) cases represented 15% (11-23%) of the 71 000 (range 47000 – 94000) pulmonary TB patients estimated to have MDR-TB  in 2013 in this Region.  The reported cases  among retreatment cases represented x% of the 18000 (range 14000 – 21000) estimated MDR-TB among retreatment cases. Following countries detected 60% of the estimated cases among retreatment:  COUNTRIES and %.........  The reported MDR-TB (R resistant) among new cases represented x% of the 57000 (range 40000 – 74000) estimated MDR-TB among new.

Among notified cases X number of cases are below 15 years of age. Male female ratio of reported cases is X: X.

All the cases detected were not enrolled on treatment. MDR-TB cases started on second-line treatment for MDR-TB in 2013 was 6921 which is 65% of the notified cases. There was steady increase in case enrolment over the years which are prominent from 2011 to 2014. However the ratio of the numbers of patients starting treatment for MDR-Tb, to those notified was also increased. The gap pf notified and enrolled are growing in recent years in China, Philippines and Viet Nam.

<a id="f.alignment"></a> 

##### `r I(pasteLabel("Figure", figCount, "f.alignment", insLink=FALSE))`. Trend of notification and enrolment of MDR-TB cases in the Western Pacific Region, 2006`r paste0("\u2013", yr-2)`
Graph showing notification and enrolment
``` {r f.alignment, fig.width=9, fig.height=9}
f.alignment
```

<sup>a</sup> Only countries reporting both indicators are included in the WPR aggregate.

#### Treatment outcome
``` {r d.tx.out}
txo1 <- subset(mdat, year==yr-2 & g_whoregion=="WPR" & !is.na(mdr_coh) & mdr_coh!=0, select=c(country, mdr_coh, mdr_succ, mdr_def, mdr_died))
txo1$c_tsrm <- rounder(txo1$mdr_succ / txo1$mdr_coh * 100)
txo2 <- nrow(txo1)
txo3 <- rounder(sum(txo1$mdr_succ) / sum(txo1$mdr_coh) * 100)
txo4 <- rounder(sum(txo1$mdr_def) / sum(txo1$mdr_coh) * 100)
txo5 <- rounder(sum(txo1$mdr_died) / sum(txo1$mdr_coh) * 100)
txo6 <- txo1[txo1$c_tsrm == max(txo1$c_tsrm), "country"]
txo7 <- txo1[txo1$c_tsrm == min(txo1$c_tsrm), "country"]
```

`r txo2` countries reported their treatment outcome of 2011 cohort. In the 2011 cohort, the proportion of MDR-TB patients who successfully completed treatment was `r txo3`%, loss to follow up was `r txo4`% and died `r txo5`%. Treatment success was highest in `r txo6` and lowest in `r txo7`. The treatment success in China is consistently low over years, death and failure rate remained high. Malaysia and Philippines showed continuous fall of success rate. Non evaluation of cases is the main cause in Malaysia while loss to follow up play major role in decreasing trend of success rate in the Philippines. Among high TB burden countries Cambodia and Viet Nam shows good outcome which remains over 70% over the years.


DO WE KNOW OUTCOME OF XDR_TB COHORT IN ANY COUNTRY?

<a id="f.tx.out"></a> 

##### `r I(pasteLabel("Figure", figCount, "f.tx.out", insLink=FALSE))`.  Trend of treatment outcome expressed as a proportion among MDR-TB cases in the Western Pacific Region, 2006--`r yr-2`
``` {r f.tx.out, fig.width=9, fig.height=7}
f.tx.out
```

			

### XDR-TB

Data on resistance to FQ and injectable agents among cases with MDR-TB were reported from 12 countries. Among notified MDR-TB cases 38% cases had their DST on 2nd line Tb drugs. Combining data from all countries, a total of 7% of patients with MDR-TB had resistance to either a FQ or a second line injectable agent or both, the proportion of MDR-TB cases with resistance to FQ is 4% and with resistance to second line injectable agents is 3%.  2% of MDR-TB cases are XDR-TB. XDR-TB had been reported by 6 countries in the Region (17% of all countries and areas in the Region) in 2013. Total reported XDR-TB cases are 106 (2% of MDR-TB cases). The highest rate is shown in Republic of Korea which is 9% of all MDR-TB cases.

<a id="t.xdr"></a> 

##### `r I(pasteLabel("Table", tableCount, "t.xdr", insLink=FALSE))`.			Rate (FQ and injectable)
```{r t.xdr, results='asis'}
t.xdr
```

### Funding situation; drug cost

X million USD was spent for MDR-TB management which was X% of the total budget of NTP in 2013. Among the total spent X% is from domestic source (Government allocation), X% from loan and X% from grant. Second line drugs are X% of the total cost for MDR-TB management. Expenditure for SLD has been significantly increased over years.


<a id="f.expend"></a> 

##### `r I(pasteLabel("Figure", figCount, "f.expend", insLink=FALSE))`. Trend of notification and enrolment of MDR-TB cases in the Western Pacific Region, 2006`r paste0("\u2013", yr-2)`
Graph showing notification and enrolment
``` {r f.expend, fig.width=9, fig.height=9}
f.expend
```

### WPRO response
In 2011 the Western Pacific Region Green Light Committee (WPR rGLC) was established to assist member countries with responsible scale-up of programmatic management of drug resistant TB. The rGLC initiative shifted the focus from “regulation and control” to “support and assistance”.  The WPR rGLC is composed of a secretariat based at the regional WHO office in Manila, and an expert committee with complimentary skill sets. Key activities included PMDT evaluation, critical review of mission reports, contributions to regional capacity building and strategic and technical guidance.  The expert committee had regular teleconferences and bi-annual face-to-face meetings.  The Secretariat organized monitoring missions to all high TB burden countries and arranged critical review of all country reports within strict timelines.  Priority recommendations were clearly communicated and progress reviewed during subsequent visits. Regional training of trainers, consultant courses on PMDT and drug management were organized, as well as in-country training on clinical aspects of MDR-TB management.  Recognizing the important role of the private sector within the region, rGLC assisted selected countries to assess the scope of private sector involvement in PMDT.  rGLC also worked with countries regarding the introduction and protection of novel TB drugs. A regional workshop focusing on strengthening  and harmonizing the regulation of TB medicines was organized where drug regulatory authority and national TB control programme of high TB burden countries discussed various topics including registration, quality assurance, rational use of drugs and pharmacovigilence and developed priority actions. 	

	

Discussion
---------------------------------------------------



The region included countries with high MDR-TB caseloads such as China, Viet Nam and the Philippines and Pacific Island nations with very small, irregular caseloads. Member States of the Western Pacific Region are also at different stages of PMDT implementation
National representative drug resistance survey data is available from 5 out of 7 high TB burden countries of the Region (Cambodia, China, Mongolia, Philippines and Viet Nam). Among them Cambodia, Mongolia, Philippines and Viet Nam have two national representative survey data point and China is conducting it's 2nd national DRS. Two remaining countries without representative data are PNG and Lao PDR. Subnational DRS in PNG is ongoing and results are expected to be available in 2015. Surveillance data are available from 19 countries and areas of the Region. Countries need to strengthen their routine surveillance system at least for all previously treated cases. This is the 1st step of universal access of MDR-Tb diagnosis.

A PARAGRAPH ON DATA INCOMPLETENESS >>>>>>>>

Several countries such as Viet Nam and the Philippines showed commendable progress. These countries are quickly moving to nationwide expansion. The PMDT program in China has continued to evolve. To scale up PMDT, the first step is to establish diagnostic capacity. Introduction of rapid diagnostic (XPERT MTB/RIF) has paved the path. It is evident that countries has started using Xpert for diagnosis. In Viet Nam 81% of notified MDR-TB cases are from Xpert. However testing coverage remained low. In the Global plan to stop TB (REf…) 100% of retreatment cases and 20% of new cases were set as a target. Current coverage for retreatment cases is 21%. It should be noted that some countries did not report Xpert result and Rifampicin resistant cases. There is a sign of increasing trend in notification and enrolment of MDR-TB cases, so far the scale-up of the programmatic management of MDR-TB (PMDT) is slow with only 15% of the estimated MDR-TB among notified cases being reported under PMDT. As a logical step, most of the countries start PMDT targeting retreatment cases first. Data showed that countries have made tremendous progress to identify among retreatment cases (X% of estimated retreatment cases are notified). Countries still need to plan to increase the testing coverage among all retreatment cases.  Further scaling up to find MDR-TB cases among the new cases is a huge challenge. The strategy needs to be reviewed considering current financial and human resource capacity. Identifying high risk group groups among new Tb cases may be the reasonable way forward. To identify the high risk groups, surveillance data need to be strengthened.
 
Detection of MDR-Tb cases is challenging and needs lots of effort. But if the notified cases are not put on treatment, all the efforts will be in vain. For prevention and control of MDR-TB this is paramount important that all patients with reported MDR-TB are put on treatment. It is alarming that 35% of the notified cases are not put on treatment in 2013. This is 27% in China, 43% in Philippines and 36% in Viet Nam. There may be two types of situation- some cases can be in the waiting list because of unavailability of treatment capacity including drugs and another reason could be loss to follow up of confirmed cases and failed to put them on treatment. In both situations, these patients are back in community and continue to spread MDR-TB strain. Proper planning needs to be done to ensure availability of drugs and treatment provision. Political commitment to scale up treatment capacity matching the diagnostic capacity will be essential. But for the later cause, initial loss to follow up, underlying cause need to be identified and addressed. Most of the time weak health system is responsible for losing the patient initially.  Accessibility of PMDT centers and substantial out of pocket expenditure play major role. The call for integrated patient centered TB care and prevention and bold policies and supportive system in the post 2015 TB strategy need to be operationalised to address this issue. Only PMDT or NTP may not be able to address this mis alignment. 
  
 
Treatment success rates are low, below 50%. The treatment success rate in Region is similar as the global average of the treatment success rate.With this kind of low success rate, it will be hard to keep faith on PMDT with current policy. The high number of loss to follow up and non evaluated acses (X% are loss to follow up and X% are not evaluated) are responsible for this low success rate and spreading TB strain with extensive drug resistance. High proportion of loss to follow up is seen in the Philippines and high proportion of non evaluated cases are seen in Malaysia. But there are some good country examples. Cambodia and Viet Nam showed treatment success above 70%.  

WPR rGLC oversight ensured optimal value extraction from country missions.  All reports were critically reviewed with clear formulation of feasible recommendations.  Bottlenecks identified include the complexity of MDR-TB management and the need for a broader health systems approach with development of workforce, infra-structure, referral and financing mechanisms in the context of universal health care. A compelling finding was the diversity of needs at different stages of PMDT implementation in various countries, requiring different types of technical assistance. rGLC created a platform for sharing experiences, both the frustrations and successes, strengthening local ownership and encouraging innovative local solutions.  
The local ownership at the TB program level has improved, however domestic resources allocated to PMDT remains inadequate, demonstrating the need for greatly increased advocacy and strong political commitment to contain the spread of MDR-TB. While the long-term sustainability of programs was threatened by donor dependence. Donor-driven targets may also have unintended consequences, such as fragmenting service delivery and encouraging inaccurate reporting to save face. The delicate balance between ambition, vision, reality, feasibility and measurability has to be considered at all times.

The increasing trend in detection and enrolment of MDR-TB cases demonstrates readiness for PMDT scale-up at the country level. However, challenges remain considerable with serious concern regarding the long-term sustainability of donor-funded programs. MDR-TB is a man made phenomenon and unless we address the root cause it may be difficult to address MDR-TB. Prevention of MDR-TB including strengthening basic TB control needs to be at the center of the strategy to address MDR-TB. Ensuring scaling up of PMDT by increased DST coverage, ensuring all notified cases are on treatment, ensuring all enrolled patient complete their treatment need to be emphasized simultaneously.  TB control has entered the most dynamic phase since decades with many opportunities.  We have new diagnostic tools, new TB drugs and new strategies. It is therefore critical to invest both financial and technical resources to prevent MDR TB and scale up programmatic management of drug resistantTB. 
 
### Conflicts of interest
None declared.

### Funding
None.


References
--------------------------------------------------------




