---
title: "Drug resistant tuberculosis in the WHO Western Pacific Region"
author: "Tauhid Islam,<sup>&dagger;</sup> Tom Hiatt<sup>&dagger;</sup> and Nobuyuki
  Nishikiori<sup>&dagger;</sup>"
output:
  html_document:
    fig_caption: yes
    number_sections: no
    theme: united
    toc: yes
    keep_md: true
  word_document:
    fig_height: 7
---


```{r setup, echo=FALSE, include=FALSE}

###################################################
# Tom Hiatt
# Updated: 12 March 2014
# Regional MDR-TB annual analysis
###################################################

## Change this information to customize for your own data.

# Year of TB notification data
yr <- 2013
# 

###################################################

# > sessionInfo()

require("reshape")
require("ggplot2")
require("grid")
require("scales")
require("xtable")
require("stringr")
require("timeSeries")
require("ggthemes")
require("Gmisc")


# Reproducible research process inspired by: http://gforge.se/2014/01/fast-track-publishing-using-knitr-part-ii/ and http://gforge.se/2014/07/fast-track-publishing-using-rmarkdown/

##################
# Knitr settings #
##################
require(knitr) 
knitr::opts_chunk$set(warning=FALSE,
    			  message=FALSE,
					  echo=FALSE,
					  dpi=96,
					  fig.width=4, fig.height=4, # Default figure widths
					  dev="png", dev.args=list(type="cairo"), # The png device
					  # Change to dev="postscript" if you want the EPS-files
					  # for submitting. Also remove the dev.args() as the postscript
					  # doesn't accept the type="cairo" argument.
					  error=FALSE)

# Create regional report folder if needed.
# dir.create("./Regional_report") # run this if error below
# dir.create("./figure data") # run this if error below
if(!any(grep("MDR_report", getwd()))) setwd("./MDR_report")

# Theme for plots
theme_report <- function(base_size=12, base_family="") {
    colors <- ggthemes_data$few
    gray <- colors$medium['gray']
    black <- colors$dark['black'] # I don't know why these last 3 parts are needed, but they are. List is here: http://docs.ggplot2.org/0.9.2.1/theme.html
    theme_bw(base_size=base_size, base_family=base_family) +
        theme(
            line = element_line(colour = gray),
            rect = element_rect(fill = "white", colour = NA),
            text = element_text(colour = black),
            axis.ticks.x = element_line(colour = gray),
            axis.ticks.y = element_blank(),
            legend.key = element_rect(colour = NA),
            ## Examples do not use grid lines
            panel.border = element_rect(colour = gray),
            panel.grid.major.x = element_blank(),
            panel.grid.minor = element_blank(),
            strip.background = element_rect(fill="white", colour=NA),
            strip.text = element_text(hjust=0)
            )
}

########################################################
# Functions for data formatting
########################################################

# Simple rounder that just adds in the thousands separator 
rounder <- function(x, decimals=FALSE) {
  if(decimals==TRUE){
    ifelse(is.na(x), NA, ifelse(x==0, 0, ifelse(x < 0.01, "<0.01", ifelse(round(x,2) < 0.1, formatC(round(x,2), format='f', digits=2), ifelse(round(x,1) < 10, formatC(round(x,1), format='f', digits=1), formatC(round(x,0), big.mark=" ", format='d') )))))
  }
  else ifelse(is.na(x), NA, ifelse(x==0, 0, ifelse(x < 1, "< 1", formatC(round(x,0), big.mark=" ", format='d'))))
}

.rowsums <- function(x) { 
  # This function sums rows ignoring NAs unless all are NA
  # use it like this
  # t3c$snu <- .rowsums(t3c[c('new_sn', 'new_su')])
  	tosum <- as.matrix(x)
		summed <- rowMeans((tosum), na.rm=T) * rowSums(!is.na((tosum)))
		return(summed)
}

# Change names to WPSAR convention
.WPSARnames <- function(d, col='country', ord='no order'){
  d[col] <- as.character(d[[col]])
  d[col] <- ifelse(d[[col]]=='China, Hong Kong SAR', 'Hong Kong Special Administrative Region (China)', 
                   ifelse(d[[col]]=='China, Macao SAR', 'Macao Special Administrative Region (China)', 
                          ifelse(d[[col]]=='Micronesia (Federated States of)', 'Micronesia, Federated States of', 
                                 ifelse(d[[col]]=='WPR', 'Western Pacific Region
', d[[col]]))))

  if(!ord %in% c('wpr')) warning('Not ordering.')
if(ord=='wpr')  d <- d[match(c("Afghanistan", "Bangladesh", "Brazil", "Cambodia", "China", "DR Congo", "Ethiopia", "India", "Indonesia", "Kenya", "Mozambique",  "Myanmar", "Nigeria", "Pakistan", "Philippines", "Russian Federation", "South Africa", "Thailand", "Uganda", "UR Tanzania", "Viet Nam", "Zimbabwe", "High-burden countries", "AFR", "AMR", "EMR", "EUR", "SEAR", "WPR", "Global"), d[[col]]),]

  return(d)
}

########################################################
# Functions for maps
########################################################

source("./MapFunctions.r")
load("gparts.Rdata")

########################################################
# Functions to assist with tables and figure in markdown document (from here: http://rmflight.github.io/posts/2012/10/papersinRmd.html)
########################################################

incCount <- function(inObj, useName) {
  nObj <- length(inObj)
  useNum <- max(inObj) + 1
  inObj <- c(inObj, useNum)
  names(inObj)[nObj + 1] <- useName
  inObj
}
figCount <- c(`_` = 0)
tableCount <- c(`_` = 0)

tableCount

pasteLabel <- function(preText, inObj, objName, insLink = TRUE, sepper = " ") {
  objNum <- inObj[objName]
  
  useText <- paste(preText, objNum, sep = sepper)
  if (insLink) {
    useText <- paste("[", useText, "](#", objName, ")", sep = " ")
  }
  useText
}

tableCat <- function(inFrame) {
  outText <- paste(names(inFrame), collapse = " | ")
  outText <- c(outText, paste(rep("---", ncol(inFrame)), collapse = " | "))
  invisible(apply(inFrame, 1, function(inRow) {
    outText <<- c(outText, paste(inRow, collapse = " | "))
  }))
  return(outText)
}

################################################
# Create directory and Get data 
################################################

external.data <- FALSE

if(!external.data){
  source("D:/Users/hiattt/Dropbox/Code/R/.Rprofile")
  runprofile()
} 

if(external.data){
 if(!"data" %in% dir() | length(dir("./data"))==0){
  dir.create(paste("./data"))
  stop("Download the notification and treatment outcomes data from 'http://who.int/tb/country/data/download/en/' to './data'.")
  }

data1 <- dir("./data", full.names=TRUE)
tb <- NA
for(i in data1){
  data3 <- read.csv(i)
  tb <- merge(tb, data3, all=TRUE)
}

# Note: to use your own data, make a flat data file with each row corresponding to the lower reporting unit and a aggregating variable. EG by country with aggregating variable of region.

# Add needed variables
tb$g_hbc22 <- ifelse(tb$iso3 %in% c("AFG", "BGD", "BRA", "CHN", "COD", "ETH", "IDN", "IND", "KEN", "KHM", "MMR", "MOZ", "NGA", "PAK", "PHL", "RUS", "THA", "TZA", "UGA", "VNM", "ZAF", "ZWE"), "high", "low")

tb$c_new <- .rowsums(tb[c("new_sp", "new_sn", "new_su", "new_ep", "new_oth")])
tb$c_newinc <- .rowsums(tb[c("c_new", "ret_rel", "newret_oth")])
tb$c_ret <- .rowsums(tb[c("ret_rel", "ret_taf", "ret_tad", "ret_oth")])
tb$c_notified <- .rowsums(tb[c("c_new", "c_ret", "newret_oth")])

tb$c_tot_newrel_100k <- tb$tot_newrel / tb$e_pop_num * 1e5

# Get population data (available from UN population division. http://esa.un.org/unpd/wpp/unpp/panel_population.htm) and formatted TB/HIV data and estimates for the WHO Western Pacific Region.

} # End of external data part


# Common data munging 

whbc <- subset(tb, g_whoregion=="WPR" & year==yr & mdr>=1, c(iso3))
whbc <- whbc$iso3
               
#                c("KHM", "CHN", "JPN", "MYS", "MNG", "PNG", "PHL", "KOR", "VNM")

tb$country2 <- ifelse(tb$iso3 %in% whbc, as.character(tb$iso3), "Other") 
tb$country2 <- factor(tb$country2, c("KHM", "CHN", "JPN", "MYS", "MNG", "PNG", "PHL", "KOR", "VNM", "Other"), c("Cambodia", "China", "Japan", "Malaysia", "Mongolia", "Papua New Guinea", "Philippines", "Republic of Korea", "Viet Nam", "Other countries"))


```



<sup>&dagger;</sup> Stop TB and Leprosy Elimination, Division of Combating Communicable Diseases, World Health Organization Regional Office for the Western Pacific, Manila, Philippines.

Correspondence to Tauhid Islam (e-mail: islamt@wpro.who.int).

Running title: 

Key words: 

Abstract 
--------------------------------------------------------




Introduction
--------------------------------------------------------




Methods
--------------------------------------------------------

[modify this text appropriately.]

### Data
Every year, 36 countries and areas in the Region are requested to report TB surveillance data to WHO using a standardized data collection form. Since 2009, a web-based online system has been used for data submission and validation. Collected data covers the following areas: TB case notifications and treatment outcomes, diagnostic and treatment services, drug management, surveillance and surveys of drug-resistance, information on TB/HIV co-infection, infection control, engagement of all care providers and budgets and expenditures for TB control. The full description of methods is available in the Global Tuberculosis Report 2013 and the data sets are available from the WHO global TB database (www.who.int/tb/data). Case definitions for TB can be found in the 4th edition of the TB treatment guidelines.^4 In 2013, 30 countries and areas of the Western Pacific Region reported data representing more than 99.9% of the total population. This report described the epidemiological situation and progress in programmatic response with a focus on seven countries with a high burden of TB: Cambodia, China, the Lao People's Democratic Republic, Mongolia, Papua New Guinea, the Philippines and Viet Nam. (Globally, WHO designates 22 countries with a high burden of TB that include Cambodia, China, the Philippines and Viet Nam. The other three countries: the Lao People's Democratic Republic, Mongolia and Papua New Guinea are considered priority countries with a high burden of TB in the Western Pacific Region).

```{r Versions}
info <- sessionInfo()
r_ver <- paste(info$R.version$major, info$R.version$minor, sep=".")
```
 
All analyses were performed using R (ver. `r r_ver`)[R Core Team, 2013] and packages rms (ver. `r info$otherPkgs$rms$Version`) [F. Harrell, 2014] for analysis, Gmisc for plot and table output (ver. `r info$otherPkgs$Gmisc$Version`), and knitr (ver `r info$otherPkgs$knitr$Version`) [Xie, 2013] for reproducible research.
 

### Analysis and reproducibility
Analysis was conducted by the statistical package R (R Core Team, 2013, Vienna, Austria, www.R-project.org). Due to calls for transparent and reproducible research,^5,6 we have published programme code to generate the entire contents of this article including all figures and tables by using R with the knitr package (Yihui Xie, 2013). Readers can download the code (see supplement material) and reproduce all figures and tables under an appropriate personal computing environment. For non-commercial purposes, readers may modify the code to produce figures and tables that are not presented in this article. For instance, readers may wish to produce tables and figures for countries or regions other than the WHO Western Pacific Region.


Results
--------------------------------------------------------

```{r m-coverage-data, warning=FALSE}
figCount <- incCount(figCount, "m-coverage")

coa <- read.table(textConnection("
iso3 coverage year
AUS Nationwide	2013
BRN	Nationwide	2013
KHM	Nationwide	2007
CHN	Nationwide	2014
HKG	Nationwide	2012
MAC	Nationwide	2013
COK	Nationwide	2013
FJI	Nationwide	2006
PYF	Nationwide	2013
GUM	Nationwide	2012
JPN	Nationwide	2002
MYS	Subnational	1997
MHL	Nationwide	2013
MNG	Nationwide	2007
NCL	Nationwide	2012
NZL	Nationwide	2012
MNP	Nationwide	2012
PLW	Nationwide	2013
PNG	Subnational	2014
PHL	Nationwide	2012
KOR	Nationwide	2004
WSM	Nationwide	2013
SGP	Nationwide	2013
SLB	Nationwide	2004
VUT	Nationwide	2006
VNM	Nationwide	2012
"), header = TRUE, as.is = TRUE)
# closeAllConnections() # This breaks stuff if left in there.

coa$cat <- cut(coa$year, c(0, 2005, 2010, 2014, Inf), c('1997–2004', '2005–2009', '2010–2013', 'Ongoing survey in 2014'), right=FALSE)
# 
mc1 <- WHOmap.print(coa, legend.title= "Year of most \nrecent data", copyright=FALSE, show=FALSE, zoom="WPR")

write.csv(coa, file=paste0(pasteLabel("./figure_data/table", tableCount, "m-coverage", insLink=FALSE, sepper=""), ".csv"), row.names=FALSE, na="")

```

### Surveillance
#### Coverage of drug resistance surveillance: 

<a id="m-coverage"></a> 

##### `r I(pasteLabel("Figure", figCount, "m-coverage", insLink=FALSE))`.	Progress in regional coverage of surveillance data on drug resistance, `r paste0(range(coa$year), collapse="\u2013")`<sup>a</sup>

``` {r m-coverage, fig.width=9, fig.height=9}

mc1

```

<sup>a</sup> Coverage in Malaysia and Papua New Guinea is at the subnational level only.



```{r t-est-enroll-data, warning=FALSE}
tableCount <- incCount(tableCount, "t-est-enroll")

# Notification table

oa <- subset(tb, year==yr & g_whoregion=="WPR", 
              select=c(iso3, g_whoregion, country, 
                       e_mdr_num, e_mdr_num_lo, e_mdr_num_hi,
                       mdr, rapid_dx_dr_r, 
                       conf_mdr_tx, unconf_mdr_tx))

oa <- oa[order(oa$country),]
names(oa)[names(oa)=='country'] <- 'area'
estvars <- c("e_mdr_num", "e_mdr_num_lo", "e_mdr_num_hi")

# make aggregate rows
ob <- oa[oa$iso3 %in% whbc, ]
obh <- aggregate(ob[4:ncol(ob)], by=list(area=ob$g_whoregion), FUN=sum, na.rm=TRUE)
obh$area <- "Total"
obh[estvars] <- NA # Not the proper way to sum bounds. Maybe to replace this later.

obw <- aggregate(oa[4:ncol(oa)], by=list(area=oa$g_whoregion), FUN=sum, na.rm=TRUE)
obw[estvars] <- subset(emdra, year==yr & group_name=="WPR", estvars) # Proper sums for aggregates.

# combine together
oc <- rbind(ob[3:ncol(ob)], obh, obw) 

# calculate and format vars

oc$e_mdr_num_range <- paste0("(", rounder(oc$e_mdr_num_lo), "&#8211;", rounder(oc$e_mdr_num_hi), ")")

oc$pmdr <- .rowsums(oc[c('mdr', 'rapid_dx_dr_r')])
oc$mdr_tx <- .rowsums(oc[c('conf_mdr_tx', 'unconf_mdr_tx')])

oc$pmdr_pct <- rounder(oc$pmdr / oc$e_mdr_num * 100)
oc$pmdr_pct_range <- paste0("(", rounder(oc$pmdr / oc$e_mdr_num_hi * 100), "&#8211;", rounder(oc$pmdr / oc$e_mdr_num_lo * 100), ")")

oc$mdr_tx_pct <- rounder(oc$mdr_tx / oc$pmdr * 100)

# Format
oc$e_mdr_num <- rounder(oc$e_mdr_num)

for(var in 2:ncol(oc)){
  oc[var] <- ifelse(is.na(oc[[var]]), "&#8211;", oc[[var]])
}


# Rename countries
od <- .WPSARnames(oc, col="area")

rownames(od) <- od$area

oe <- subset(od, select=c("e_mdr_num", "e_mdr_num_range", "pmdr", "pmdr_pct", "pmdr_pct_range", "mdr_tx", "mdr_tx_pct"))


write.csv(oe, file=paste0(pasteLabel("./figure_data/table", tableCount, "t-est-enroll", insLink=FALSE, sepper=""), ".csv"), row.names=FALSE, na="")

```

<a id="t-est-enroll"></a> 

##### `r I(pasteLabel("Table", tableCount, "t-est-enroll", insLink=FALSE))`. Estimated drug resistant cases compared with notified cases and cases put on treatment, `r yr`
```{r t-est-enroll, results='asis'}
             
# Make that table
htmlTable(oe, caption = "", rowlabel = "", cgroup = rbind(c("MDR-TB cases<sup>*</sup>", "Enrolled on treatment", rep(NA,3)), c("Estimated", "Detected", "%", "Number", "%")), n.cgroup = rbind(c(3,2, rep(NA,3)), c(2,1,2,1,1)), align=c(rep(c('r','l','r'),2), 'r'), ctable = TRUE, tfoot = "<sup>*</sup> All columns except estimates include Rif-resistant cases confirmed by Xpert only.", headings = NA )

```


```{r t-drnotif-data, warning=FALSE}
tableCount <- incCount(tableCount, "t-drnotif")

# Notification table

nb <- subset(tb, year==yr & g_whoregion=="WPR", 
              select=c(iso3, g_whoregion, country, 
                       cullpa_new, cullpa_ret, cullpa_unk, 
                       dst_rlt_new, dst_rlt_ret, dst_rlt_unk, 
                       dr_h_nr_new,  dr_h_nr_ret,  dr_h_nr_unk, 
                       dr_r_nh_new,  dr_r_nh_ret,  dr_r_nh_unk, 
                       mdr_new,  mdr_ret,  mdr_unk,
                       xpert_new, xpert_ret, xpert_unk, 
                       xpert_dr_r_new, xpert_dr_r_ret, xpert_dr_r_unk))

nb <- nb[order(nb$country),]
names(nb)[names(nb)=='country'] <- 'area'

# make aggregate rows
nbc <- nb[nb$iso3 %in% whbc, ]
nbch <- aggregate(nbc[4:ncol(nbc)], by=list(area=nbc$g_whoregion), FUN=sum, na.rm=TRUE)
nbch$area <- "Total"

nbcw <- aggregate(nb[4:ncol(nb)], by=list(area=nb$g_whoregion), FUN=sum, na.rm=TRUE)

# combine together
nbd <- rbind(nbc[3:ncol(nbc)], nbch, nbcw) # , tbbh, tbbga

# calculate and format vars
nbd$cullpa <- .rowsums(nbd[c('cullpa_new', 'cullpa_ret', 'cullpa_unk')])
nbd$dst_rlt_new_pct <- nbd$dst_rlt_new / nbd$cullpa_new * 100
nbd$dst_rlt_ret_pct <- nbd$dst_rlt_ret / nbd$cullpa_ret * 100
nbd$dst_rlt <- .rowsums(nbd[c('dst_rlt_new', 'dst_rlt_ret', 'dst_rlt_unk')])
nbd$dst_rlt_pct <- nbd$dst_rlt /  nbd$cullpa * 100

nbd$dr_h_nr_new_pct <- nbd$dr_h_nr_new / nbd$dst_rlt_new * 100
nbd$dr_h_nr_ret_pct <- nbd$dr_h_nr_ret / nbd$dst_rlt_ret * 100
nbd$dr_h_nr_pct <- .rowsums(nbd[c('dr_h_nr_new', 'dr_h_nr_ret', 'dr_h_nr_unk')]) /  nbd$dst_rlt * 100

nbd$dr_r_nh_new_pct <- nbd$dr_r_nh_new / nbd$dst_rlt_new * 100
nbd$dr_r_nh_ret_pct <- nbd$dr_r_nh_ret / nbd$dst_rlt_ret * 100
nbd$dr_r_nh_pct <- .rowsums(nbd[c('dr_r_nh_new', 'dr_r_nh_ret', 'dr_r_nh_unk')]) /  nbd$dst_rlt * 100

nbd$mdr_new_pct <- nbd$mdr_new / nbd$dst_rlt_new * 100
nbd$mdr_ret_pct <- nbd$mdr_ret / nbd$dst_rlt_ret * 100
nbd$mdr_pct <- .rowsums(nbd[c('mdr_new', 'mdr_ret', 'mdr_unk')]) /  nbd$dst_rlt * 100

nbd$xpert <- .rowsums(nbd[c('xpert_new', 'xpert_ret', 'xpert_unk')])

nbd$xpert_dr_r_new_pct <- nbd$xpert_dr_r_new / nbd$xpert_new * 100
nbd$xpert_dr_r_ret_pct <- nbd$xpert_dr_r_ret / nbd$xpert_ret * 100
nbd$xpert_dr_r_pct <- .rowsums(nbd[c('xpert_dr_r_new', 'xpert_dr_r_ret', 'xpert_dr_r_unk')]) /  nbd$xpert * 100

# Format
for(var in 2:ncol(nbd)){
  nbd[var] <- ifelse(is.na(nbd[[var]]), "&#8211;", rounder(nbd[[var]]))
}


# Rename countries
nbe <- .WPSARnames(nbd, col="area")

# Consider adding a column for put on treatment and/or including those from the other MDR table.

rownames(nbe) <- nbe$area

nbf <- subset(nbe, select=c("cullpa_new", "cullpa_ret", "cullpa", "dst_rlt_new_pct", "dst_rlt_ret_pct", "dst_rlt_pct", "dr_h_nr_new_pct", "dr_h_nr_ret_pct", "dr_h_nr_pct", "dr_r_nh_new_pct", "dr_r_nh_ret_pct", "dr_r_nh_pct", "mdr_new_pct", "mdr_ret_pct", "mdr_pct", "xpert_new", "xpert_ret", "xpert", "xpert_dr_r_new_pct", "xpert_dr_r_ret_pct", "xpert_dr_r_pct"))

names(nbf) <- c("New", "Ret.", "Total", "New", "Ret.", "Total", "New", "Ret.", "Total", "New", "Ret.", "Total", "New", "Ret.", "Total", "New", "Ret.", "Total", "New", "Ret.", "Total")



write.csv(nbf, file=paste0(pasteLabel("./figure_data/table", tableCount, "t-drnotif", insLink=FALSE, sepper=""), ".csv"), row.names=FALSE, na="")

```


#### MDR-TB rate:

<a id="t-notif"></a> 

##### `r I(pasteLabel("Table", tableCount, "t-notif", insLink=FALSE))`. Table: rate among new and reTx, CI, year, number of country year data point

(R resistance data need to be included as separate column)

<a id="t-drnotif"></a> 

##### `r I(pasteLabel("Table", tableCount, "t-drnotif", insLink=FALSE))`. Drug resistance in select countries, `r yr`


```{r t-drnotif, results='asis'}

# Make that table
htmlTable(nbf, caption = "", rowlabel = "", cgroup = c("Cases with culture or LPA results", "% with DST results", "% with resistance to H only", "% with resistance to R only", "% MDR-TB<sup>*</sup>", "Additional cases confirmed by Xpert", "% Xpert cases MDR-TB<sup>*</sup>"), n.cgroup = c(3,3,3,3,3,3,3), ctable = TRUE, tfoot = "<sup>*</sup> Resistant to H and R. Totals also include History Unknown.")
```


			

#### XDR-TB:
<a id="t-notif"></a> 
##### `r I(pasteLabel("Table", tableCount, "t-notif", insLink=FALSE))`.			Table, number and year of reporting

```{r t-xdr-data, warning=FALSE}
tableCount <- incCount(tableCount, "t-xdr")

# Notification table

xa <- subset(tb, year==yr & g_whoregion=="WPR", 
              select=c(iso3, g_whoregion, country, 
                       mdr, mdr_dst_rlt, mdr_ds_fq2li,
                       mdr_dr_fq, mdr_dr_2li, xdr))

xa <- xa[order(xa$country),]
names(xa)[names(xa)=='country'] <- 'area'

# make aggregate rows
xb <- xa[xa$iso3 %in% whbc, ]
xc <- aggregate(xb[4:ncol(xb)], by=list(area=xb$g_whoregion), FUN=sum, na.rm=TRUE)
xc$area <- "Total"

xd <- aggregate(xa[4:ncol(xa)], by=list(area=xa$g_whoregion), FUN=sum, na.rm=TRUE)

# combine together
xe <- rbind(xb[3:ncol(xb)], xc, xd) # , tbbh, tbbga

# calculate and format vars
xe$xdst_pct <- xe$mdr_dst_rlt / xe$mdr * 100
xe$fq_pct <- xe$mdr_dr_fq / xe$mdr * 100
xe$xli_pct <- xe$mdr_dr_2li / xe$mdr * 100
xe$xdr_pct <- xe$xdr / xe$mdr * 100

# Format
for(var in 2:ncol(xe)){
  xe[var] <- ifelse(is.na(xe[[var]]), "&#8211;", rounder(xe[[var]]))
}


# Rename countries
xf <- .WPSARnames(xe, col="area")

rownames(xf) <- xf$area

xg <- subset(xf, select=c("mdr", "xdst_pct", "mdr_dr_fq", "fq_pct", "mdr_dr_2li", "xli_pct", "xdr", "xdr_pct"))

names(xg) <- c("MDR-TB", "% 2<sup>nd</sup>-line DST", rep(c("Number", "%"),3))

write.csv(xf, file=paste0(pasteLabel("./figure_data/table", tableCount, "t-drnotif", insLink=FALSE, sepper=""), ".csv"), row.names=FALSE, na="")

```

#### Resistance to SLD: 
<a id="t-xdr"></a> 

##### `r I(pasteLabel("Table", tableCount, "t-xdr", insLink=FALSE))`.			Rate (FQ and injectable)


```{r t-xdr, results='asis'}

# Make that table
htmlTable(xg, caption = "", rowlabel = "", cgroup = c("", "Resistance to FQ", "Resistance to 2<sup>nd</sup>-line injectible", "XDR-TB"), n.cgroup = c(2,2,2,2), ctable = TRUE, tfoot = "")
```



#### Trends in drug resistance:
<a id="t-notif"></a> 
##### `r I(pasteLabel("Table", tableCount, "t-notif", insLink=FALSE))`.			Possible?? Graph?
	
### Management of drug resistant TB
		
#### DST coverage
<a id="t-notif"></a> 
##### `r I(pasteLabel("Table", tableCount, "t-notif", insLink=FALSE))`.			Table : % of DST result among new and reTx
			
#### Notification and enrolment
<a id="t-notif"></a> 
##### `r I(pasteLabel("Table", tableCount, "t-notif", insLink=FALSE))`.			Year, number notified and enrolled, % of total burden
			Age, sex? Child MDR-TB?

#### Alignment


```{r f-align-bar-data}

figCount <- incCount(figCount, "f-align-bar")

tea <- subset(tb, year %in% 2007:yr & g_whoregion=="WPR", select=c(iso3, country, year, e_mdr_num, e_mdr_num_lo, e_mdr_num_hi, mdr, rapid_dx_dr_r, conf_mdr_tx, unconf_mdr_tx))

teb1 <- aggregate(tea[4:ncol(tea)], by=list(year=tea$year), FUN=sum, na.rm=TRUE)
teb1$country <- teb1$iso3 <- "WPR"

# Separate hbcs
teb3 <- tea[tea$iso3 %in% whbc,] 
# teb2 <- aggregate(teb3[4:ncol(tea)], by=list(year=teb3$year, area=teb3$country), FUN=sum, na.rm=TRUE)


#combine
teb <- rbind(teb1, teb3)

# Calculate new vars
teb$`Cases confirmed` <- .rowsums(teb[c('mdr', 'rapid_dx_dr_r')])
teb$`Patients enrolled on treatment` <- .rowsums(teb[c('conf_mdr_tx', 'unconf_mdr_tx')])


tec <- melt(teb[c("country", "year", "Cases confirmed", "Patients enrolled on treatment")], id=1:2)

ted <- ggplot(tec, aes(year, value, color=variable)) + geom_point(alpha=.5) + geom_line(size=1, alpha=.5) + facet_wrap(~country, scales="free_y", ncol=4) + theme_report() + scale_color_brewer("", type="qual", palette=6) + scale_x_continuous("") + scale_y_continuous("MDR-TB cases")  + guides(fill = guide_legend(reverse = TRUE)) + theme(legend.position = "bottom")

write.csv(tec, file=paste0(pasteLabel("./figure_data/figure", figCount, "f-align-bar", insLink=FALSE), ".csv"), row.names=FALSE)
```


<a id="f-align"></a> 

##### `r I(pasteLabel("Figure", figCount, "f-align-bar", insLink=FALSE))`. Trend of notification and enrolment of MDR-TB cases in the Western Pacific Region, 2006`r paste0("\u2013", yr-2)`
Graph showing notification and enrolment
``` {r f-align-bar, fig.width=9, fig.height=9}

ted

```



#### Treatment outcomes

```{r f-mtxout-bar-data}

figCount <- incCount(figCount, "f-mtxout-bar")

tra <- subset(tb, year %in% 2006:(yr-2) & g_whoregion=="WPR", select=c(iso3, country, year, mdr_coh, mdr_cur, mdr_cmplt, mdr_died, mdr_fail, mdr_def, mdr_succ))

# Fill in 0s for NAs
tra[is.na(tra)] <- 0

trb1 <- aggregate(tra[4:ncol(tra)], by=list(year=tra$year), FUN=sum, na.rm=TRUE)
trb1$area <- trb1$iso3 <- "WPR"

# Separate hbcs
trb2 <- subset(tra, iso3 %in% whbc & mdr_coh > 0) 
trb2 <- rename(trb2, c(country="area"))
# trb2 <- aggregate(tra[4:ncol(tra)], by=list(year=tra$year, area=tra$country), FUN=sum, na.rm=TRUE)

#combine
trb <- rbind(trb1, trb2)

trb$mdr_succ <- ifelse(trb$year>=2011, trb$mdr_succ, trb$mdr_cur + trb$mdr_cmplt)
trb$Success <- trb$mdr_succ / trb$mdr_coh * 100
trb$Died <- trb$mdr_died / trb$mdr_coh * 100
trb$Failed <- trb$mdr_fail / trb$mdr_coh * 100
trb$Defaulted <- trb$mdr_def / trb$mdr_coh * 100
trb$`Not evaluated` <- (trb$mdr_coh - (trb$mdr_succ + trb$mdr_died + trb$mdr_fail + trb$mdr_def)) / trb$mdr_coh * 100

trc <- melt(trb[c("area", "year", "Success", "Died", "Failed", "Defaulted", "Not evaluated")], id=1:2)

write.csv(trc, file=paste0(pasteLabel("./figure_data/figure", figCount, "f-mtxout-bar", insLink=FALSE), ".csv"), row.names=FALSE)

```

<a id="f-mtxout-bar"></a> 

##### `r I(pasteLabel("Figure", figCount, "f-mtxout-bar", insLink=FALSE))`.	Trend of treatment outcome expressed as a proportion among MDR-TB cases in the Western Pacific Region, 2006--`r yr-2`
``` {r f-mtxout-bar, fig.width=9, fig.height=7}

ggplot(trc[trc$area=="WPR",], aes(year, value, fill=variable)) + geom_bar(stat="identity", position="stack") + theme_report() + scale_fill_brewer('Outcome', type="qual", palette=6) + scale_x_continuous("", breaks=min(trc$year):max(trc$year)) + scale_y_continuous("Percent of cohort") + coord_cartesian(ylim=c(0,100)) + guides(fill = guide_legend(reverse = TRUE))

ggplot(trc, aes(year, value, fill=variable)) + geom_bar(stat="identity", position="stack") + facet_wrap(~area) + theme_report() + scale_fill_brewer('Outcome', type="qual", palette=6) + scale_x_continuous("") + scale_y_continuous("Percent of cohort") + coord_cartesian(ylim=c(0,100)) + guides(fill = guide_legend(reverse = TRUE))


```



Discussion
---------------------------------------------------


```{r t-drestnotif-data, warning=FALSE}
tableCount <- incCount(tableCount, "t-drestnotif")

# Notification table

oa <- subset(tb, year==yr & g_whoregion=="WPR", 
              select=c(iso3, g_whoregion, country, 
                       e_new_mdr_prop, e_new_mdr_prop_lo, e_new_mdr_prop_hi,
                       e_ret_mdr_prop, e_ret_mdr_prop_lo, e_ret_mdr_prop_hi,
                       e_mdr_num, e_mdr_num_lo, e_mdr_num_hi,
                       c_newunk, c_ret, 
                       dst_rlt_new, dst_rlt_ret, dst_rlt_unk, 
                       mdr_new,  mdr_ret,  mdr, rapid_dx_dr_r, 
                       conf_mdr_tx, unconf_mdr_tx))

oa <- oa[order(oa$country),]
names(oa)[names(oa)=='country'] <- 'area'
estvars <- c("e_new_mdr_prop", "e_new_mdr_prop_lo", "e_new_mdr_prop_hi", "e_ret_mdr_prop", "e_ret_mdr_prop_lo", "e_ret_mdr_prop_hi", "e_mdr_num", "e_mdr_num_lo", "e_mdr_num_hi")

# make aggregate rows
ob <- oa[oa$iso3 %in% whbc, ]
obh <- aggregate(ob[4:ncol(ob)], by=list(area=ob$g_whoregion), FUN=sum, na.rm=TRUE)
obh$area <- "Total"
obh[estvars] <- NA # Not the proper way to sum bounds. Maybe to replace this later.

obw <- aggregate(oa[4:ncol(oa)], by=list(area=oa$g_whoregion), FUN=sum, na.rm=TRUE)
obw[estvars] <- subset(emdra, year==yr & group_name=="WPR", estvars) # Proper sums for aggregates.

# combine together
oc <- rbind(ob[3:ncol(ob)], obh, obw) 

# calculate and format vars

for(stima in c("new", "ret")){ # change proportional estimates to percent
  for(bound in c("", "_lo", "_hi")){
    oc[paste0("e_", stima, "_mdr_pct", bound)] <- oc[paste0("e_", stima, "_mdr_prop", bound)] * 100
  }
}

oc$e_new_mdr_pct_range <- paste0("(", rounder(oc$e_new_mdr_pct_lo), "&#8211;", rounder(oc$e_new_mdr_pct_hi), ")")
oc$e_ret_mdr_pct_range <- paste0("(", rounder(oc$e_ret_mdr_pct_lo), "&#8211;", rounder(oc$e_ret_mdr_pct_hi), ")")
oc$e_mdr_num_range <- paste0("(", rounder(oc$e_mdr_num_lo), "&#8211;", rounder(oc$e_mdr_num_hi), ")")

oc$dst_rlt_new_pct <- rounder(oc$dst_rlt_new / oc$c_newunk * 100)
oc$dst_rlt_ret_pct <- rounder(oc$dst_rlt_ret / oc$c_ret * 100)

oc$mdr_new_pct <- rounder(oc$mdr_new / oc$c_newunk * 100)
oc$mdr_ret_pct <- rounder(oc$mdr_ret / oc$c_ret * 100)

oc$mdr_tx <- .rowsums(oc[c('conf_mdr_tx', 'unconf_mdr_tx')])

# Format
for(var in c("e_new_mdr_pct", "e_ret_mdr_pct", "e_mdr_num", "mdr", "rapid_dx_dr_r", "conf_mdr_tx", "mdr_tx")){
  oc[var] <- ifelse(is.na(oc[[var]]), "&#8211;", rounder(oc[[var]]))
}
for(var in c("mdr_new_pct", "mdr_ret_pct")){
  oc[var] <- ifelse(is.na(oc[[var]]), "&#8211;", oc[[var]])
}

# Rename countries
od <- .WPSARnames(oc, col="area")

rownames(od) <- od$area

oe <- subset(od, select=c("mdr_new_pct", "e_new_mdr_pct", "e_new_mdr_pct_range", "mdr_ret_pct", "e_ret_mdr_pct", "e_ret_mdr_pct_range", "mdr", "e_mdr_num", "e_mdr_num_range", "rapid_dx_dr_r", "conf_mdr_tx", "mdr_tx"))


write.csv(oe, file=paste0(pasteLabel("./figure_data/table", tableCount, "t-drestnotif", insLink=FALSE, sepper=""), ".csv"), row.names=FALSE, na="")

```

<a id="t-drestnotif"></a> 

##### `r I(pasteLabel("Table", tableCount, "t-drestnotif", insLink=FALSE))`. Estimated drug resistant cases compared with notified cases and cases put on treatment, `r yr`
```{r t-drestnotif, results='asis'}
             
# Make that table
htmlTable(oe, caption = "", rowlabel = "", cgroup = rbind(c("% MDR-TB among new", "% MDR-TB among ret.", "MDR-TB cases", "", rep(NA,5)), c(rep(c("Detected", "Estimated"),2), "Detected", "Estimated <br>___________________", "RR cases<sup>*</sup>", "Confirmed MDR-TB on treatment", "Total MDR-TB on treatment")), n.cgroup = rbind(c(3,3,3,3, rep(NA,5)), c(1,2,1,2,1,2,1,1,1)), align=c(rep(c('c','r', 'l'),3), rep("c",3) ), ctable = TRUE, tfoot = "<sup>*</sup> Additional Rifampicine resistant cases detected by Xpert.", headings = NA )

```


### WPRO response
		
#### rGLC and  activities


### Issues & way forward


Conflicts of interest
--------------------------------------------------------
None declared.

Funding
--------------------------------------------------------
None.

References
--------------------------------------------------------

1. Global Tuberculosis Report 2013. Geneva, World Health Organization, 2013 (http://www.who.int/tb/publications/global_report/en/, accessed 15 March 2014). 



